On the Functionality of the Multilayer Perceptron

### 2020-09-11 Thesis/Sections/FeatureSelections.tex
11 Sept 2020
- Updates to NN section
- Updates to feature section

\paragraph*{}We concatenate each analysis frame, $a^{(i)}, i \in [0,k-1]$ into a single $k \times N$ matrix, called $A$. Each row is a frame, each column is an index in that frame
\begin{equation}
\label{eqn-FrameMatrix}
A = \big\{ a^{(0)} , a^{(1)} , a^{(2)} , ... , a^{(k-1)} \big\} = 
\begin{bmatrix}
a^{(0)}[0] & a^{(0)}[1] & a^{(0)}[2] & \hdots & a^{(0)}[N-1] \\
a^{(1)}[0] & a^{(1)}[1] & a^{(1)}[2] & \hdots & a^{(1)}[N-1] \\
a^{(2)}[0] & a^{(2)}[1] & a^{(2)}[2] & \hdots & a^{(2)}[N-1] \\
\vdots 		& \vdots 	  & \vdots 		& \ddots & \vdots        \\
a^{(k-1)}[0] & a^{(k-1)}[1] & a^{(k-1)}[2] & \hdots & a^{(k-1)}[N-1] \\
\end{bmatrix}
\end{equation}
We use bracket notation, $[j]$ to indicate that each object is array-like. Many programming languages could also use the following indexing conventions for matrix 
$A$:
\begin{equation}
\label{eqn-Indexing}
A_{i,j} = a^{(i)}[j] = A[i][j] = A[i,j]
\end{equation}
Note that these all represent equivalent entries.

@@

\paragraph*{}Most standard implementations of neural network require all activations to be real floating-point numbers. Since the the DFT matrix introduce complex values, we compute the square of the element-wise $L2$-norm of the resultant array.
\begin{equation}
\label{eqn-Spectrogram}
S_{xx} = \big{\lVert} \mathbb{W}\widetilde{A}^T \big{\rVert}_{2}^{2}
\end{equation}
Where $\mathbb{W}$ is the DFT matrix from eqn. (\ref{eqn-DFTMatrix}) and $\widetilde{A}^T$ is the transpose of the analysis frames matrix from eqn. (\ref{eqn-WindowMatrix}). The matrix $S_{xx}$, is the spectrogram representation of the initial waveform, and now contains $N \times k$ real floating-point numbers.

\paragraph*{}Each Column of the $S_{xx}$ matrix is now a single-frame that has been moved into a frequency-space representation, thus there are $k$ columns, just as there were $k$ analysis frames. Given the discretized nature of digital audio, the fequency-sapce representation is not a continuous function, but rather a column vector, where the frequency has been assigned to one of $N$ bins. Since human hearing extends from about $20$ Hz to around $20$ kHz, audio files, such as music and voice recordings are typically sampled around $44.1$ kHz, to ensure that the full range of audio is held by the recording \cite{Olson,Virtanen}.

\paragraph*{}However, standard musical instruments seldom extend above $8$ kHz \cite{Olson,Virtanen,White}. This means that when constructing the spectrogram, we will rarely ever see energy present above this frequency at any time, and the $S_{xx}$ matrix will contain mostly zero, or zero-like entries. As a result of this, we can simply crop each matrix, to display frequencies that are only between $0$ Hz and $8000$ Hz. This makes the input array smaller, and eliminates redundant and non-useful information. This brings the number rows in the $S_{xx}$ matrix from $N$ down to $N'$. lastly, each raw audio file may contain a different number of samples to begin with, thus a different number of analysis frames are created. To ensure all sample are a homogeneous size, we assert that each matrix must have exactly $k'$ columns. If $k' < k$, columns of zeros are added to pad the input, and if $k' > k$, columns are removed.

\paragraph*{}Each spectrogram is now $N' \times k' \times 1$ ($1$ representing a single gray-scale channel, as oppose to $3$ for RGB inputs) and effectively encodes the energy distribution of the waveform a a function of both time and frequency. The spectrogram is the first feature used in this model. For this classifier, we have chosen $N' = 560$ and $k' = 256$. For training, a mini-batch of $b$ samples are concatenated into a single array object. \textcolor{blue}{See Neural Network section for more details}. For a batch of $b$ samples of $N' \times k' \times 1$ spectrograms, we shape $X_1$ such that:
\begin{equation}
\label{eqn-X1 shape}
X_1 = \big\{ S_{xx}^{(0)},S_{xx}^{(1)},S_{xx}^{(2)}, ... , S_{xx}^{(b-1)} \big\} \in \mathbb{R}^{(b \times N' \times k' \times 1)}
\end{equation}
This matrix is presented to the \textit{Convolution} branch of the neural network from processing. \textcolor{blue}{See Neural Network section for more details}

@@

\paragraph*{}The time domain envelope (TDE) is a method of determining which parts of the sound wave contains most of the energy of the signal. Generally, we start by computing the RMS-Energy of each frame. The larger the RMS energy of the frame, the larger the amplitude of the waveform in that frame. For frames with smaller amplitude, we conclude that the signal has either not begun, or is decaying. The RMS-Energy of a single analysis-frame $a^{(i)}$ is given by \cite{Olson,Virtanen}
\begin{equation}
\label{eqn-RMS}
RMS\big[ a^{(i)} \big] = \sqrt{\frac{1}{N} \sum_{j=0}^{N-1}a^{(i)}[j]}
\end{equation}

@@

\paragraph*{}The zero crossing rate (ZXR) of a signal or frame is use to measure how many times that a signal crosses it's equilibrium point. This can be done per total sound wave, or per unit time, such an per analysis-frame. In differentiating speech and music, it is commonly to use the difference between the frame with the highest ZXR and the lowest ZXR. In some cases, the ZXR can be correlated to frequency as well \cite{Kahn,Zhang}. The ZXR for a frame $a^{(i)}$ is given by \cite{Virtanen,Liu}
\begin{equation}
\label{eqn-ZXR}
ZXR\big[ a^{(i)} \big] = \frac{1}{2} \sum_{j=1}^{N-1} \big{|} \sign(a^{(i)}[j]) - \sign(a^{(i)}[j-1]) \big{|} 
\end{equation}
Where $\sign(x)$ returns $+1$ if $x > 0$, $-1$ if $x < 0$ and $0$ if $x = 0$.

@@

\paragraph*{}The temporal center of mass (TCM) of a signal is used to compute where in time the amplitude sort of \textit{bunches up}. We treat the full waveform array, $a$, with $M$ samples as a 1-Dimensional discrete mass distribution. The TCM of that waveform is then given:
\begin{equation}
\label{eqn-FeatureTCM}
TCM\big[ a \big] = \frac{\sum_{j=0}^{M-1}j a[j]}{\sum_{j=0}^{M-1}a[j]}
\end{equation}
This includes negative amplitude values acting as "negative masses".

@@

\subsubsection{Statistical Distribution Data}

@@

\paragraph*{}Auto correlation coefficients (ACC) are rough estimates of the signal spectral distribution \cite{Virtanen}. We can compute any number of ACC's and their value changed depending on the index chosen. It is common to only compute the first $k$ ACC's. For a full waveform signal $a$, with $M$ samples, the $k$-th ACC (indexed from $1$ to $K$) is given by:
\begin{equation}
\label{eqn-FeatureACC}
ACC_k\big[ a \big] = \frac{\sum_{j=0}^{M-k-1}a[j]a[j+k]} {\sqrt{\sum_{j=0}^{M-k-1}a^2[j]} \sqrt{\sum_{j=0}^{M-k-1}a^2[j+k]}}
\end{equation}

@@

\paragraph*{}The features described in this section are derived from the frequency-domain representations of each audio sample. Typically, these features are exacted from the spectrogram representation,$S_xx$, outlined in sec. (\ref{subsec-spectrogram}) and computed in eqn.(\ref{eqn-Spectrogram}). For each feature, we detail the physical significance and provide a visualization in feature-space.

\subsubsection{Mel Frequency Ceptral Coefficients}

@@

\paragraph*{}The frequency center-of-mass ($f_{CoM}$) for an audio file provides a strong representation of how overtones and energy is distributed in the signal. For example, a low instrument such as a tuba or bass may have a very low $f_{CoM}$ while flutes and violins may have a very high $f_{CoM}$.

>>>>

\paragraph*{}We concatenate each analysis frame, $a^{(i)}, i \in [0,k-1]$ into a single $k \times N$ matrix, called $A$. Each row is a frame, each column is an index in that frame
\begin{equation}
\label{eqn-FrameMatrix}
A = \big\{ a^{(0)} , a^{(1)} , a^{(2)} , ... , a^{(k-1)} \big\} = 
\begin{bmatrix}
a^{(0)}[0] & a^{(0)}[1] & a^{(0)}[2] & \hdots & a^{(0)}[N-1] \\
a^{(1)}[0] & a^{(1)}[1] & a^{(1)}[2] & \hdots & a^{(1)}[N-1] \\
a^{(2)}[0] & a^{(2)}[1] & a^{(2)}[2] & \hdots & a^{(2)}[N-1] \\
\vdots 		& \vdots 	  & \vdots 		& \ddots & \vdots        \\
a^{(k-1)}[0] & a^{(k-1)}[1] & a^{(k-1)}[2] & \hdots & a^{(k-1)}[N-1] \\
\end{bmatrix}
\end{equation}
We use bracket notation, $[j]$ to indicate that each object is array-like. Many programming languages could also use the following indexing conventions for matrix 
$A$:
\begin{equation}
\label{eqn-IndexingA}
A_{i,j} = a^{(i)}[j] = A[i][j] = A[i,j]
\end{equation}
Note that these all represent equivalent entries.

@@

\paragraph*{}Most standard implementations of neural network require all activations to be real floating-point numbers. Since the the DFT matrix introduce complex values, we compute the square of the element-wise $L2$-norm of the resultant array.
\begin{equation}
\label{eqn-Spectrogram}
S = \big{\lVert} \mathbb{W}\widetilde{A}^T \big{\rVert}_{2}^{2}
\end{equation}
Where $\mathbb{W}$ is the DFT matrix from eqn. (\ref{eqn-DFTMatrix}) and $\widetilde{A}^T$ is the transpose of the analysis frames matrix from eqn. (\ref{eqn-WindowMatrix}). The matrix $S$, is the spectrogram representation of the initial waveform, and now contains $N \times k$ real floating-point numbers. We can index matrix $S$ similarly to that of matrix $A$ in eqn. (\ref{eqn-IndexingA}):
\begin{equation}
\label{eqn-IndexingS}
S_{i,j} = s^{(i)}[j] = S[i][j] = S[i,j]
\end{equation}

\paragraph*{}Each Column of the $S$ matrix is now a single-frame that has been moved into a frequency-space representation, thus there are $k$ columns, just as there were $k$ analysis frames. Given the discretized nature of digital audio, the frequency-space representation is not a continuous function, but rather a column vector, where the frequency has been assigned to one of $N$ bins. Since human hearing extends from about $20$ Hz to around $20$ kHz, audio files, such as music and voice recordings are typically sampled around $44.1$ kHz, to ensure that the full range of audio is held by the recording \cite{Olson,Virtanen}.

\paragraph*{}However, standard musical instruments seldom extend above $8$ kHz \cite{Olson,Virtanen,White}. This means that when constructing the spectrogram, we will rarely ever see energy present above this frequency at any time, and the $S$ matrix will contain mostly zero, or zero-like entries. As a result of this, we can simply crop each matrix, to display frequencies that are only between $0$ Hz and $8000$ Hz. This makes the input array smaller, and eliminates redundant and non-useful information. The number rows in the $S$ matrix is reduced from $N$ down to $N'$. lastly, each raw audio file may contain a different number of samples to begin with, thus a different number of analysis frames are created. To ensure all sample are a homogeneous size, we assert that each matrix must have exactly $k'$ columns. If $k' < k$, columns of zeros are added to pad the input, and if $k' > k$, columns are removed.

\paragraph*{}Each spectrogram is now $N' \times k' \times 1$ ($1$ representing a single gray-scale pixel channel, as opposed to $3$ channels for RGB inputs) and effectively encodes the energy distribution of the waveform a a function of both time and frequency. The spectrogram is the first feature used in this model. For this classifier, we have chosen $N' = 560$ and $k' = 256$. For training, a mini-batch of $b$ samples are concatenated into a single array object. \textcolor{blue}{See Neural Network section for more details}. For a batch of $b$ samples of $N' \times k' \times 1$ spectrograms, we shape $X_1$ such that:
\begin{equation}
\label{eqn-X1 shape}
X_1 = \big\{ S^{(0)},S^{(1)},S^{(2)}, ... , S^{(b-1)} \big\} \in \mathbb{R}^{(b \times N' \times k' \times 1)}
\end{equation}
This matrix is presented to the \textit{Convolution} branch of the neural network from processing. \textcolor{blue}{See Neural Network section for more details}

@@

\paragraph*{}The time domain envelope (TDE) is a method of determining which parts of the sound wave contains most of the energy of the signal. Generally, we start by computing the RMS-Energy of each frame. The larger the RMS energy of the frame, the larger the amplitude of the waveform in that frame. For frames with smaller amplitude, we conclude that the signal has either not begun, or is decaying. The RMS-Energy of a single analysis-frame $a^{(i)}$ is given by \cite{Olson,Virtanen}
\begin{equation}
\label{eqn-RMS}
\text{RMS}\big[ a^{(i)} \big] = \sqrt{\frac{1}{N} \sum_{j=0}^{N-1}a^{(i)}[j]}
\end{equation}

@@

\paragraph*{}The zero crossing rate (ZXR) of a signal or frame is use to measure how many times that a signal crosses it's equilibrium point. This can be done per total sound wave, or per unit time, such an per analysis-frame. In differentiating speech and music, it is commonly to use the difference between the frame with the highest ZXR and the lowest ZXR. In some cases, the ZXR can be correlated to frequency as well \cite{Kahn,Zhang}. The ZXR for a frame $a^{(i)}$ is given by \cite{Virtanen,Liu}
\begin{equation}
\label{eqn-ZXR}
\text{ZXR}\big[ a^{(i)} \big] = \frac{1}{2} \sum_{j=1}^{N-1} \big{|} \sign(a^{(i)}[j]) - \sign(a^{(i)}[j-1]) \big{|} 
\end{equation}
Where $\sign(x)$ returns $+1$ if $x > 0$, $-1$ if $x < 0$ and $0$ if $x = 0$.

@@

\paragraph*{}The temporal center of mass (TCM) of a signal is used to compute where in time the amplitude sort of \textit{bunches up}. We treat the full waveform array, $a$, with $M$ samples as a 1-Dimensional discrete mass distribution. The TCM of that waveform is then given:
\begin{equation}
\label{eqn-FeatureTCM}
\text{TCM}\big[ a \big] = \frac{\sum_{j=0}^{M-1}j a[j]}{\sum_{j=0}^{M-1}a[j]}
\end{equation}
This includes negative amplitude values acting as "negative masses".

@@

\subsubsection{Statistical Distribution Data}
\textcolor{red}{Honetly, I may remove these from the classifier}

@@

\paragraph*{}Auto correlation coefficients (ACC) are rough estimates of the signal spectral distribution \cite{Virtanen}. We can compute any number of ACC's and their value changed depending on the index chosen. It is common to only compute the first $k$ ACC's. For a full waveform signal $a$, with $M$ samples, the $k$-th ACC (indexed from $1$ to $K$) is given by:
\begin{equation}
\label{eqn-FeatureACC}
\text{ACC}_k\big[ a \big] = \frac{\sum_{j=0}^{M-k-1}a[j]a[j+k]} {\sqrt{\sum_{j=0}^{M-k-1}a^2[j]} \sqrt{\sum_{j=0}^{M-k-1}a^2[j+k]}}
\end{equation}

@@

\paragraph*{}The features described in this section are derived from the frequency-domain representations of each audio sample. Each of these features are exacted from the transpose spectrogram representation, $S^T$, where the creation of $S$ is outlined in sec. (\ref{subsec-spectrogram}) and computed in eqn.(\ref{eqn-Spectrogram}). This is done to return analysis frame into a row-like array, representing frequency space divided into $N'$ bins, with $k'$ analysis frames. For each feature, we detail the physical significance and provide a visualization in feature-space.

\subsubsection{Mel Frequency Cepstral Coefficients}

\paragraph*{}Mel Frequency Ceptstral Coefficients (MFCC's) are a very prolific feature used widly in audio classification tasks \cite{Liu,Virtanen,Zhang}. MFCC's are produced by converting a linear-frequency axis with units such as Hertz into units of Mels, which are design to account for the non-linear perception of frequency in humans. The Hertz to Mel and Mel to Hertz transforms are provided \cite{Virtanen}:
\begin{equation}
\label{eqn-HztoMel}
M_f[h] = 2595 \log_{10}\big(1+ \frac{h}{700}\big)
\end{equation}
\begin{equation}
\label{eqn-MeltoHz}
H_f[m] = 700 \big(10^{(\frac{m}{2595})}-1\big)
\end{equation}
Where $M_f$ is the frequency in units of Mels, given $[h]$, a frequency in Hertz, and $H_f$ is the frequency in Hertz given $[m]$ a frequency in Mels.

\paragraph*{}Mel filters are produced by grouping frequency-space into $R$ overlapping bins called \textit{Mel Filter-banks}. In units of Mels, each filter bank, $R_i$ remains a constant width, but when shown in Hertz, the bins increase logarithmically. 
\begin{figure}[H]
\label{fig-MelFilterBanks}
\begin{center}
\textcolor{red}{Figure of a few Mel Filter Banks goes here}
\end{center}
\caption{Mel Filter Banks shown in frequency space with units of Hertz}
\end{figure}
Each of the $R$ filters is created to be $N'$ samples long, and transformed back into units of Hertz. When applied to an analysis frame in the frequency spectrum, the dot-product between the filter and the spectrum gives an approximation of the energy in that filter bank. Thus, each filter is concatenated into a matrix $M$ of shape $R \times N'$, where each row is a filter. We apply the Mel Filter banks to the spectrogram to create matrix $B$:
\begin{equation}
\label{eqn-FilterBanks}
B = S^T M^T
\end{equation}
Matrix $B$ has shape $k' \times R$.

\paragraph*{}Using $S^T$ ensure that each row of the matrix is an analysis frame in frequency space. Similarly, each column of the matrix $M$ is a appropriately scaled Mel filter-bank. This way, the matrix product in eqn. (\ref{eqn-FilterBanks}) allows that $B_{i,j}$ is the dot product between the $i$-th analysis frame and the $j$-th filter-bank. Finally, we compute the average energy across all $k'$ frames. The resulting $R$ floating-point numbers are approximation of the signal's energy in those particular frequency bands. For this project, we have chosen to use $R = 12$ filter-banks, which are all used as features.

\begin{figure}[H]
\label{fig-FeatureMFCC}
\begin{center}
\textcolor{red}{PLOT OF MFCC FEATURES HERE}
\end{center}
\caption{Mel Frequency Ceptral Coefficients in feature-space}
\end{figure}

@@

\paragraph*{}The frequency center-of-mass (FCM) for an audio file provides a strong representation of how overtones and energy is distributed in the signal's frequency domain. As with the temporal center-of-mass, we treat each row of the $S$ matrix as it's own 1-Dimensional mass distribution, and compute the center of mass of that row. This encodes the FCM for a single analysis frame (in frequency-space) in the waveform. For an frequency-analysis frame, $s^{(i)}$, the FCM is given by:
\begin{equation}
\label{eqn-FeatureFCM}
\text{TCM}_{i}\big[ s^{(i)} \big] = \frac{\sum_{j=0}^{N'-1}j s^{(i)}[j]}{\sum_{j=0}^{N'-1}s^{(i)}[j]}
\end{equation}
We compute the FCM for each of the $k'$ frames, and then average the results. We use the average FCM across $k'$ frames to compute the FCM feature:
\begin{equation}
TCM = \sum_{i=0}^{k'} \text{TCM}_{i}\big[ s^{(i)} \big]
\end{equation}

\paragraph*{}The average FCM gives a strong approximation of the instrument or signal source's range. For example, a flute or violin will have a considerably high FCM value, even in their lower registers. Similarly, basses or tubas will have considerably low FCM values. Given that the standard frequency range of some musical instruments is fixed, it is guaranteed that for any particular instrument, the FCM will consistently remain within certain bounds \cite{Olson,White}.

\begin{figure}[H]
\label{fig-FeatureFCM}
\begin{center}
\textcolor{red}{PLOT OF FCM FEATURES HERE}
\end{center}
\caption{Frequency center-of-mass visualized in feature-space}
\end{figure}

\subsection{The Design Matrix}
